---
title: "Exploring the impact of closed areas on trawl survey biomass index trends"
output: html_document
---

Let's load the survey data frames to work with. These are generated from the PBSsynopsis package. We will work with 4 example species.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
theme_set(ggsidekick::theme_sleek())
```

```{r load_dat, cache=TRUE}
areas <- readRDS("../pbs-synopsis/data-cache/stratum-areas.rds")

survey_tows <- readRDS("../pbs-synopsis/data-cache/all-survey-spatial-tows.rds") %>%
  inner_join(areas, by = c("survey_id", "grouping_code"))

spp <- c("shortraker rockfish", "lingcod", "dover sole", "canary rockfish",
  "pacific ocean perch")
survey_tows_select <- filter(survey_tows, species_common_name %in% spp)
```

First we need to write a function that calculates biomass for a given year and survey by taking the mean density within each stratum and summing those densities multiplied by the stratum areas:

```{r calc_bio}
calc_bio <- function(dat, i = seq_len(nrow(dat))) {
  dat[i, ] %>% group_by(year, survey_id, area_km2, grouping_code) %>% 
    summarise(density = mean(density_kgpm2*1e6)) %>% 
    group_by(year) %>% 
    summarise(biomass = sum(density * area_km2)) %>% 
    pull(biomass)
}
```

Let's test our function against the previously calculated indices. The red line represents the existing index. That grey line represents our own calculation.

```{r bio, fig.width=5, fig.height=3.5}
bio <- survey_tows %>% 
  filter(species_common_name %in% "pacific ocean perch") %>%
  filter(survey_series_id == 3) %>% 
  calc_bio()

boot_ind <- readRDS("../pbs-synopsis/data-cache/all-boot-biomass-indices.rds") %>%
  filter(species_common_name %in% "pacific ocean perch") %>% 
  filter(survey_series_id == 3)

ggplot(data.frame(biomass = bio, year = boot_ind$year),
  aes(year, biomass)) + geom_line(lwd = 4, col = "grey80") +
  geom_line(data = boot_ind, colour = "red") +
  ylim(0, NA)
```

They are exactly the same. So our biomass calculation function works.

Now let's pipe our data through a bootstrap function to calculate the percentile bootstrap confidence intervals. Note that the `.` below refers to the respective chunk of data.

```{r boot_biomass}
boot_biomass <- function(dat, reps = 100) {
  out <- dat %>%
    group_by(year, species_common_name, survey_series_desc) %>% 
    do({
      b <- boot::boot(., statistic = calc_bio, strata = .$grouping_code, R = reps)
      suppressWarnings(bci <- boot::boot.ci(b, type = "perc"))
      tibble::tibble(
        mean_boot = mean(b$t), 
        median_boot = median(b$t), 
        lwr = bci$percent[[4]], 
        upr = bci$percent[[5]], 
        cv = sd(b$t)/mean(b$t),
        biomass = calc_bio(.))
    })
}
```

```{r boot, cache=TRUE, dependson=c('calc_bio', 'load_dat', 'boot_biomass')}
out <- boot_biomass(survey_tows_select, reps = 100)
```

```{r glimpse}
glimpse(out)
```

And plot that out:

```{r multiplot, fig.width=11, fig.height=8}
ggplot(out, aes(year, biomass, fill = survey_series_desc)) + 
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.3) +
  geom_line() + geom_point() +
  facet_wrap(species_common_name~survey_series_desc, scales = "free_y", ncol = 4) +
  ylim(0, NA) + 
  guides(fill = FALSE)
```

Now let's try randomly removing 1/2 of the tows:

**(The next step is to manually draw candidate closed areas and remove the tows based on their latitude and longitude. I would expect this to have a larger effect since it removes consistent spatial blocks.)**

```{r cut1}
set.seed(42)
nrow(survey_tows_select)
survey_tows_cut1 <- survey_tows_select %>% 
  group_by(survey_series_desc) %>% 
  sample_frac(size = 1/2) %>% 
  ungroup()
nrow(survey_tows_cut1)
```

```{r boot1, cache=TRUE, dependson=c('calc_bio', 'load_dat', 'boot_biomass', 'cut1')}
out1 <- boot_biomass(survey_tows_cut1, reps = 100)
```

And plot a comparison:

```{r comparison1, fig.width=11, fig.height=8}
mutate(out, type = "all") %>% 
  bind_rows(mutate(out1, type = "restricted")) %>% 
  ggplot(aes(year, biomass, fill = type)) + 
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2) +
  geom_line(aes(colour = type)) +
  facet_wrap(species_common_name~survey_series_desc, scales = "free_y", ncol = 4) +
  ylim(0, NA) + 
  guides(fill = FALSE)
```

So it can have a considerable effect, especially for species that are caught less frequently (e.g. shortraker rockfish). Again, I imagine these effects will be more dramatic with consistent spatial sections removed.

How do the CVs compare? Thin lines represent individual years and the thick lines represent the mean CVs across years. 

```{r, fig.width=8}
means <- mutate(out, type = "all") %>% 
  bind_rows(mutate(out1, type = "restricted")) %>% 
  group_by(type, survey_series_desc, species_common_name) %>% 
  summarise(mean_cv = mean(cv, na.rm = TRUE)) %>% 
  reshape2::dcast(survey_series_desc + species_common_name ~ type, 
    value.var = "mean_cv")

all <- mutate(out, type = "all") %>% 
  bind_rows(mutate(out1, type = "restricted")) %>% 
  reshape2::dcast(year + survey_series_desc + species_common_name ~ type, 
    value.var = "cv")

ggplot(all) + 
  facet_wrap(~survey_series_desc, scales = "free_y") +
  geom_segment(aes(x = 1, xend = 2, y = all, yend = restricted, 
    colour = species_common_name), alpha = 0.3) +
  geom_segment(data = means, aes(x = 1, xend = 2, y = all, yend = restricted, 
    colour = species_common_name), lwd = 2) + 
  ylab("CV") +
  xlab("") +
  scale_x_continuous(breaks = c(1, 2), labels = c("All", "Restricted"))
  # scale_colour_brewer(palette = "Dark2")
```

This has less of an overall effect than I would have expected although year-to-year there can be fairly dramatic increases in the annual CV calculated through the bootstrap.

Someday down the road this will be a nice framework for comparing spatiotemporal model-based indexed standardization to the bootstrap approach with spatial closures.
